A simulation framework for exploring spatio-temporal dynamics in mixed fisheries

### 2018-01-19 write_up/mixfishsim_ecolmod.tex
Updated manuscript outline and added timer to sim

%% Numbered
%\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\address[1]{Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{UPMC}
\address[4]{Wageningen Marine}

\begin{abstract}
[A concise and factual abstract is required. The abstract should state briefly
the purpose of the research, the principal results and major conclusions. An
abstract is often presented separately from the article, so it must be able to
stand alone. For this reason, References should be avoided, but if essential,
then cite the author(s) and year(s). Also, non-standard or uncommon
abbreviations should be avoided, but if essential they must be defined at their
first mention in the abstract itself.  
Graphical abstract: Although a graphical abstract is optional, its use is
encouraged as it draws more attention to the online article. The graphical
abstract should summarize the contents of the article in a concise, pictorial
form designed to capture the attention of a wide readership. Graphical
abstracts should be submitted as a separate file in the online submission
system. Image size: Please provide an image with a minimum of 531 X 1328 pixels
(h X w) or proportionally more. The image should be readable at a size of 5 x
13 cm using a regular screen resolution of 96 dpi.  Preferred file types: TIFF,
EPS, PDF or MS Office files.]
\end{abstract}

@@

Fishers exploit fish populations that are heterogenously distributed without
prior knowledge of species distributions and non-selective fishing gear.
Fisheries are managed by single-species quotas, leading to discarding of
overquota catch. Increasing interest in technical solutions such as gear,
spatial closures as a way of managing "fisheries" instead of fish stocks. \\

Provide some examples here of where this is the case..focus on spatial
management and the increasing sophistication / resolution of data dervied from
VMS and logbooks; refs: Geritsen et al, Mateo et al. \\

Use of spatial management as a tool is hampered by lack of knowledge of fish
and fishery spatiotemporal dyanmics and the scale at which processes are
important for management. Need to understand these to implement effective
management measures for species avoidance / targeting. Use Dunn et al as a
ref.\\

We develop a simulation model where underlying population dynamics are known
rather than inferred from sampling or commercial catches. It includes
population movement and a number of different fisheries exploiting four fish
populations with different demographics.\\

Using the model, we simulation 20 years of exploitation of the fish populations
and use the results to draw inference on the underlying population structures -
as done in species modelling approaches. \\

@@

[State the objectives of the work and provide an adequate background, avoiding a
detailed literature survey or a summary of the results.]

@@

[Provide sufficient details to allow the work to be reproduced by an independent
researcher. Methods that are already published should be summarized, and
indicated by a reference.  If quoting directly from a previously published
method, use quotation marks and also cite the source. Any modifications to
existing methods should also be described.]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview schematic  
\begin{figure}[!ht]
\begin{tikzpicture}[%
    >=triangle 60,              % Nice arrows; your taste may be different
    start chain=going below,    % General flow is top-to-bottom
    node distance=8mm and 60mm, % Global setup of box spacing
    every join/.style={norm}   % Default linetype for connecting boxes
    ]                 
    \label{fig:model}
    % ------------------------------------------------- 
% A few box styles 
% <on chain> *and* <on grid> reduce the need for manual relative
% positioning of nodes
\tikzset{
  base/.style={draw, on chain, on grid, align=center, minimum height=4ex},
  proc/.style={base, rectangle, text width=8em},
  test/.style={base, rectangle, text width=5em},
  term/.style={proc, rounded corners},
  % coord node style is used for placing corners of connecting lines
  coord/.style={coordinate, on chain, on grid, node distance=6mm and 25mm},
  % nmark node style is used for coordinate debugging marks
  nmark/.style={draw, cyan, circle, font={\sffamily\bfseries}},
  % -------------------------------------------------
  % Connector line styles for different parts of the diagram
  norm/.style={->, draw, lcnorm},
  free/.style={->, draw, lcfree},
  cong/.style={->, draw, lccong},
  it/.style={font={\small\itshape}}
}
% -------------------------------------------------
% Start by placing the nodes in the middle
\node [proc, densely dotted, it] (p0) {start};
% Use join to connect a node to the previous one 
\node [term, join, fill=lcnorm!25]      {Choose random location to fish};
\node [term, join, fill=lcfree!25] (p1){Sample catch};
\node [test, densely dotted] (t1) {if $y > y-1$};
\node [test, join, densely dotted] (t2) {if $spp_{t}$ != $spp_{t-1}$};
\node [test, join, densely dotted] (wk) {if $week_{t}$ != $week_{t-1}$};
\node [test, join, densely dotted] (t3) {if $trip_{t}$ != $trip_{t-1}$};
\node [term, join, fill=lcnorm!25]  (p2) {Correlated Random Walk};
\node [term, join, fill=lcnorm!25]  (p3) {Choose new fishing location};
\node [proc, densely dotted, join, it] (p7) {end};

>>>>

%% Numbered
%\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\address[1]{Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{Université Pierre et Marie Curie, 4 Place Jussieu, 75005 Paris,
	France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}

\begin{abstract}
[A concise and factual abstract is required. The abstract should state briefly
the purpose of the research, the principal results and major conclusions. An
abstract is often pnresented separately from the article, so it must be able to
stand alone. For this reason, References should be avoided, but if essential,
then cite the author(s) and year(s). Also, non-standard or uncommon
abbreviations should be avoided, but if essential they must be defined at their
first mention in the abstract itself.  
Graphical abstract: Although a graphical abstract is optional, its use is
encouraged as it draws more attention to the online article. The graphical
abstract should summarize the contents of the article in a concise, pictorial
form designed to capture the attention of a wide readership. Graphical
abstracts should be submitted as a separate file in the online submission
system. Image size: Please provide an image with a minimum of 531 X 1328 pixels
(h X w) or proportionally more. The image should be readable at a size of 5 x
13 cm using a regular screen resolution of 96 dpi.  Preferred file types: TIFF,
EPS, PDF or MS Office files.]
\end{abstract}

@@

Fishers exploit fish populations that are heterogenously distributed in space
and time without prior knowledge of species distributions and non-selective
fishing gear. Fisheries managed by single-species quotas still catch an
assemblage of species, known as mixed fisheries, leading to discarding of
overquota catch. Reducing discarding is crucial to ensure biological and
economic sustainability of fisheries and as such there is increasing interest
in technical solutions such as gear, spatial closures as a way of avoiding
discarding unwanted catch. \\

Use of spatial management as a tool is hampered by lack of knowledge of fish
and fishery spatiotemporal dyanmics and understanding of the scale at which
processes are important for management, crucial in order to implement measures
at the right resolution to ensure effective management\cite{Dunn2016} which
promote species avoidance for vulnerable or low quota species while allowing
continance of sustainable fisheries for available quota species. \\

Ensuring measures are implemented at an appropriate scale has been a challenge
in the a past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective wirth increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea\cite{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data\cite{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016}, though such information is patchy and derived from a biased
sampling programme (i.e.  targeted fishing). Further, generally fishers only
recorded landings (not catch) on a daily basis. This leads to questions about the
validity of inferences that can be drawn from landings data assigned to VMS
activity pings. \\

In order to test the assumption that VMS associated landings data can be used
to draw inference on the underlying population structures we develop a
simulation model where population dynamics are known rather than inferred from
sampling or commercial catches, population movement driven by a random
(diffusive) and directed (advective) process and characterisation of a number
of different fisheries exploiting four fish populations with different spatial
and population demographics.\\

Using our model we simulation 20 years of exploitation of the fish populations
and use the results to draw inference on the underlying population structures.
We compare this inference to i) a simulated stratified fixed-site sampling
design commonly used for fisheries monitoring purposes, ii) the underlying
population structures input to the simulation.\\

[Could fit a geostatistical model (e.g. VAST) to the fisheries-dependent and
fisheries-independent data, though may be overkill...] \\

@@

[Guidance:: State the objectives of the work and provide an adequate
background, avoiding a detailed literature survey or a summary of the results.]
\\

@@

[Provide sufficient details to allow the work to be reproduced by an
independent researcher. Methods that are already published should be
summarized, and indicated by a reference.  If quoting directly from a
previously published method, use quotation marks and also cite the source. Any
modifications to existing methods should also be described.] \\

We developed a simulation model with a modular discrete-event approach to each
of the processes so that they can occur on different time-scales. The fishing
model operated on a tow-by-tow basis, while population dynamics (fishing and
natural mortality, growth) operate on a daily time-step.  Population movement
occurs on a two-weekly time-step, while recruitment occurs periodcally each
year for a set time period (e.g. 3 weeks) at at specified point individual to a
species. The model structure is summerised in Figure 1.  \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview schematic  
\begin{figure}[!ht]
	\hspace{-4em}
\begin{tikzpicture}[%
    >=triangle 60,              % Nice arrows; your taste may be different
    start chain=going below,    % General flow is top-to-bottom
    node distance=8mm and 60mm, % Global setup of box spacing
    every join/.style={norm},   % Default linetype for connecting boxes
    scale=0.9,
    every node/.style={scale=0.8}]                 
    \label{fig:model}
    % ------------------------------------------------- 
% A few box styles 
% <on chain> *and* <on grid> reduce the need for manual relative
% positioning of nodes
\tikzset{
  base/.style={draw, on chain, on grid, align=center, minimum height=4ex},
  proc/.style={base, rectangle, text width=8em},
  test/.style={base, rectangle, text width=5em},
  term/.style={proc, rounded corners},
  % coord node style is used for placing corners of connecting lines
  coord/.style={coordinate, on chain, on grid, node distance=6mm and 25mm},
  % nmark node style is used for coordinate debugging marks
  nmark/.style={draw, cyan, circle, font={\sffamily\bfseries}},
  % -------------------------------------------------
  % Connector line styles for different parts of the diagram
  norm/.style={->, draw, lcnorm},
  free/.style={->, draw, lcfree},
  cong/.style={->, draw, lccong},
  it/.style={font={\small\itshape}}
}
% -------------------------------------------------
% Start by placing the nodes in the middle
\node [proc, densely dotted, it] (p0) {start};
% Use join to connect a node to the previous one 
\node [term, join, fill=lcnorm!25]      {Choose random location to fish};
\node [term, join, fill=lcfree!25] (p1){Sample catch};
\node [test, densely dotted] (t1) {if $y > y-1$};
\node [test, join, densely dotted] (t2) {if $spp_{t}$ != $spp_{t-1}$};
\node [test, join, densely dotted] (wk) {if $week_{t}$ != $week_{t-1}$};
\node [test, join, densely dotted] (t3) {if $trip_{t}$ != $trip_{t-1}$};
\node [term, join, fill=lcnorm!25]  (p2) {Correlated Random Walk};
\node [term, join, fill=lcnorm!25]  (p3) {Choose new fishing location};
\node [proc, densely dotted, join, it] (p7) {end};
